Proof equivalence in MLL is PSPACE-complete

### 2014-12-05 main/LMCS.tex
re-worked section "Proof Nets" (minus "notation")

\cite{Girard-1987,Danos-Regnier-1989}
\cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996}
\cite{Lamarche-Strassburger-2006}
\cite{HughesFreeStar,HughesMLLProofNets}

@@

In this section, we will give a brief introduction to multiplicative linear logic (MLL). For simplicity of exposition, we will work in the unit-only fragment of MLL. The formulae of this fragment are given by the following grammar.
%
\setMidspace{8pt}
\[
	A,B,C \Coloneq \bot \Mid 1 \Mid A\parr B \Mid A\tn B
\]
%
The connectives tensor ($\tn$) and par ($\parr$) will be considered up to associativity, and \emph{duality}, indicated $\dual A$, is via DeMorgan.
%
A \emph{sequent} $\Gamma, \Delta,\Lambda$ will be a multiset of formulae.
%
To distinguish different occurrences of a formula within a sequent, the connectives and units in a sequent will be \emph{named} with distinct elements from an arbitrary set of names, e.g.:\
\[
	\named a1\named b\parr\named c1~,~\named d\bot\named e\tn\named f\bot~.
\]
%
This simple technical device is introduced for a dual purpose. Firstly, it distinguishes proofs that are equal up to a symmetric exchange of formulae, such as the \emph{identity} and the \emph{twist} proof of the above sequent, while the sequent rules can be given using standard multiset sequents, i.e.\ without the need for explicit \emph{exchange} rules. Secondly, it creates an easy way to extract proof nets, as graphs that use the names of connectives and units as vertices.

@@

The proof system we shall be working with is that consisting of just the introduction rules for the two units and the two connectives, the four rules $(\bot)$, $(\1)$, $(\parr)$, and $(\tn)$ of Figure~\ref{fig:MLL}. The problem we consider is the following.

\begin{definition}
\label{def:equivalence}
\emph{Equivalence} $(\perm)$ of proofs in (cut-free, identity-free, unit-only) multiplicative linear logic is the congruence generated by the permutations given in Figure~\ref{fig:permutations}. \emph{\capsabbrev{mll}\ proof equivalence} is the problem of deciding whether two given proofs are equivalent.
\end{definition}

@@

The permutations of sequent proofs are exactly the identifications imposed by the categorical semantics of \capsabbrev{mll}, star-autonomous categories \cite{Barr-1991} (and semi-star-autonomous categories \cite{Houston-2008,Heijltjes-Strassburger} for \capsabbrev{mll} without units).
%
Proof equivalence for \capsabbrev{mll} is therefore equivalent to the \emph{word problem} for star-autonomous categories: the problem whether two term representations denote the same morphism in any star-autonomous category.

@@

Proof nets provide a solution to proof equivalence for \capsabbrev{mll} without units.
%
For full \capsabbrev{mll}, they reduce the proof equivalence problem to a simple rewiring relation \cite{HughesMLLProofNets}.

\begin{definition}
\label{def:proof nets}
%
For a sequent $\Gamma$,
\begin{itemize}

	\item
	a \emph{linking} $\links$ is a function from the names of $\bot$-subformulae to the names of $1$-subformulae,

	\item
	a \emph{switching graph} for $\links$ is an undirected graph over the names of $\Gamma$, with for every subformula $\named aA\named c\tn\named bB$ the edges $\edge ac$ and $\edge bc$, for every subformula $\named aA\named c\parr\named bB$ either the edge $\edge ac$ or the edge $\edge bc$, and for every subformula $\named a\bot$ the edge $\edge a{\links(a)}$,

 	\item
	a \emph{proof net} $\links$ or $(\Gamma,\links)$ is a linking $\links$ such that every switching graph is acyclic and connected.

@@

\noindent
An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}.
%
The restriction that jumps must target $\1$-occurrences (rather than any connective) is a convenience---it can be circumvented by replacing a subformula $A$ by the equivalent $A\tn\1$.

@@

\begin{definition}
\label{def:proof net equivalence}
%
A \emph{rewiring} $(\perm*)$ between proof nets is the redirection of exactly one link.
%
\emph{Equivalence} $(\perm)$ of proof nets over a sequent $\Gamma$ is the equivalence generated by rewiring.
%
\end{definition}

\noindent
An example rewiring sequence is given in Figure~\ref{fig:net rewiring}, using the notation for proof nets introduced below.

To translate a sequent proof to a proof net requires to find a target for each jump from a $\bot$-formula.
%
The inference rule for $\bot$ introduces it into a sequent $\Gamma$; in the corresponding proof net, any occurrence of $\1$ in $\Gamma$ may serve as the target of the jump.

\begin{definition}
\label{def:proofs to nets}
%
The relation $(\toNet)$ interprets a proof $\Pi$ for a sequent $\Delta$ by a linking $\links$ as follows:
% 
$\Pi\toNet\links$ if for each $\named a\bot$ in $\Delta$, if $\Gamma$ is the context of the inference introducing $\named a\bot$, as illustrated below, then $\links(a)$ is the name of some $1$ in $\Gamma$.
\[
	\infer[\MLLlabel b]{\Gamma,\named a\bot}{\Gamma}
\]
%
\end{definition}

\begin{proposition}[\cite{Danos-Regnier-1989}, \cite{Lamarche-Strassburger-2006}]
\label{prop:correctness and sequentialisation}
%
%
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $\links$ is a proof net for $\Gamma$.
%
For a net $\links$ for $\Gamma$, there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (\emph{sequentialisation}).
%
\end{proposition}

\noindent
Proof nets are a canonical representation of proofs in the absence of the units: they factor out the permutations among tensor- and par-inferences, which are the last three permutations in Figure~\ref{fig:permutations}.
%
Equivalence of proof nets is generated by the four remaining permutations, on $\bot$-introduction.

\begin{proposition}[\cite{HughesMLLProofNets}]
\label{prop:proof nets work}
%
For proofs $\Pi$, $\Pi'$ and proof nets $\links$, $\links'$, if $\Pi\toNet\links$ and $\Pi'\toNet\links'$, then $\Pi\perm\Pi'$ if and only if $\links\perm\links'$.
%
\end{proposition}

\noindent
The above proposition means that \capsabbrev{mll}\ proof equivalence is the problem of deciding equivalence of proof nets.

@@

We will use a concise diagrammatic notation for sequents and proof nets.
%
The units $1$ and $\bot$ are represented by a circle~($\circ$) and a disc~($\bullet$) respectively;
%
formulae related by a tensor will be connected by edges; 
formulae related by a par will be juxtaposed, and collected in a box when a par-formula is an immediate subformula of a tensor-formula (and also for illustrative purposes).
%
For example, the following denote the same sequent:
\[
	\bot\tn\bot,\bot\tn\bot,\1,(\1\parr\1\parr\1)\tn\bot
\]
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (2,0) node[bullet] (c) {} -- (3,0) node[bullet] (d) {};
	\node[circ] (1) at (4,0) {};
	\node[circ] (2) at (5.25,0) {}; \node[circ] (3) at (6.25,0) {}; \node[circ] (4) at (7.25,0) {};
	\draw[rounded corners] (4.75,-1) rectangle (7.75,1);
	\draw (7.75,0) -- (8.75,0) node[bullet] (e) {};
\end{tikzpicture}
\]
The links of a proof net are added to the sequent as coloured arrows.
%
The following example is a proof net for the above sequent.
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[circ] (1) at (6,0) {};
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (3,0) node[bullet] (c) {} -- (4,0) node[bullet] (d) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (2.center);
			\draw[bend right=10] (b) to (3.center);
			\draw[bend left=10]  (c) to (3.center);
			\draw[bend left=10]  (d) to (4.center);
			\draw[bend right=10] (e) to (1.center);
	\end{scope}
\end{tikzpicture}
\]

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\!\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\!\parr\cdots\parr\! A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
A path in a proof net or switching graph is indicated $\path ab$, and illustrated as below.
%
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\]

@@

As the linkings above illustrate, it may occur that one subformula of a weight element $A\parr B\parr C$ fills one half of a constraint element.
%
Consequently, a weight element can fill three halves, but only of different constraint elements.
%
The other three halves may be filled by weight elements of a different edge---so to fill 3 constraint elements requires 2 inappropriate edges.
%
To fill the next three constraint elements, at most 1 previous inappropriate edge may be used, and one additional one is needed.
%
To fill $3n$ constraint elements inappropriately therefore requires $n+1$ edges.
%
It thus suffices to multiply the number of constraint elements by three times the number of edges, and encode a vertex with inflow constraint $c$ by a vertex-gadget with $c\times3\times|E|$ constraint elements (where $|E|$ denotes the number of edges in the constraint graph).

@@

In the graph of a proof net, a chosen subnet for a par can be made explicit as a \emph{box}, as illustrated below left.
%
Boxes may replace the switching condition as a correctness criterion: in the example, both the outside and the inside of the box form a tree.
%
To make this precise, we will consider the action of \emph{closing} a box, which means it is regarded as a single vertex in the graph, as illustrated below right.
%
%
%
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {}
		  (.25,0) node[bullet] (a) {} -- (1.25,0) node[bullet] (b) {}
		  ( 2.75,0) node[bullet] (c) {} -- (3.75,0) node[bullet] (d) {};
	\node[circ] (0) at (5.75,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5.5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (2);
			\draw[bend right=10] (a) to (2);
			\draw[bend right=10] (b) to (3);
			\draw[bend left=10]  (c) to (3);
			\draw[bend left=10]  (d) to (4);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\draw[dashed,rounded corners] (-.25,-.5) rectangle (4.25,4.25);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {};
	\node[circ] (0) at (3,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[draw,rounded corners,minimum size=16pt] (box) at (1,3) {};
	\draw (box)--(3,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (box);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\path (0,4.25)--(1,4.25);
\end{tikzpicture}
\]
%
%
\begin{definition}
A \emph{boxing} $s$ for a linking $\links$ for $\Gamma$ assigns a sub-sequent $s(v)\leq\Gamma$ to each par-vertex $\named v\parr$ such that 1) $v$ is a port of $s(v)$ and 2) boxes are either disjoint or strictly nested: if $s(v)\cap s(w)\neq\varnothing$ then $s(w)<s(v)$ or $s(v)<s(w)$.
\end{definition}

@@

The linearity of \capsabbrev{mll} means that in a proof or proof net, there is always a certain balance to the number of $\bot$- and $\parr$-occurrences. 
%
This observation gives a well-known necessary condition for the provability of a sequent.

>>>>

\cite{Girard-1987,Danos-Regnier-1989}
\cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996}
\cite{Koh-Ong-1999}
\cite{Lamarche-Strassburger-2006}
\cite{Hughes-2012-categories,Hughes-2012-nets}

@@

In this section, we will give a brief introduction to multiplicative linear logic (MLL). For simplicity of exposition, we will work in the unit-only fragment of MLL. The formulae of this fragment are given by the following grammar.
%
\setMidspace{8pt}
\[
	A,B,C \Coloneq \bot \Mid 1 \Mid A\pr B \Mid A\tn B
\]
%
The connectives tensor ($\tn$) and par ($\pr$) will be considered up to associativity, and \emph{duality}, indicated $\dual A$, is via DeMorgan.
%
A \emph{sequent} $\Gamma, \Delta,\Lambda$ will be a multiset of formulae.
%
To distinguish different occurrences of a formula within a sequent, the connectives and units in a sequent will be \emph{named} with distinct elements from an arbitrary set of names, e.g.:\
\[
	\named a1\named b\pr\named c1~,~\named d\bot\named e\tn\named f\bot~.
\]
%
This simple technical device is introduced for a dual purpose. Firstly, it distinguishes proofs that are equal up to a symmetric exchange of formulae, such as the \emph{identity} and the \emph{twist} proof of the above sequent, while the sequent rules can be given using standard multiset sequents, i.e.\ without the need for explicit \emph{exchange} rules. Secondly, it creates an easy way to extract proof nets, as graphs that use the names of connectives and units as vertices.

@@

The proof system we shall be working with is that consisting of just the introduction rules for the two units and the two connectives, the four rules $(\bot)$, $(\1)$, $(\pr)$, and $(\tn)$ of Figure~\ref{fig:MLL}. The problem we consider is the following.

\begin{definition}[MLL proof equivalence]
\label{def:equivalence}
\emph{Equivalence} $(\perm)$ of proofs in (cut-free, identity-free, unit-only) multiplicative linear logic is the congruence generated by the permutations given in Figure~\ref{fig:permutations}. \emph{\capsabbrev{mll}\ proof equivalence} is the problem of deciding whether two given proofs are equivalent.
\end{definition}

@@

The permutations of sequent proofs are exactly the identifications imposed by the categorical semantics of \capsabbrev{mll}, star-autonomous categories \cite{Barr-1991} (and semi-star-autonomous categories \cite{Houston-2008,Heijltjes-Strassburger} for \capsabbrev{mll} without units). Proof equivalence for \capsabbrev{mll} is therefore equivalent to the \emph{word problem} for star-autonomous categories: the problem whether two term representations denote the same morphism in any star-autonomous category.

@@

For \capsabbrev{mll} without units, Girard's original proof nets \cite{Girard-1987} are canonical: two proofs translate into the same proof net if and only if they are equivalent. Since the translation from proofs to proof nets and the syntactic comparison of proof nets are both effectively computable (linear-time), proof nets provide an effective solution to the proof equivalence problem for \capsabbrev{mll} without units.

There have been several proposals for (non-canonical) multiplicative proof nets with units \cite{Blute-Cockett-Seely-Trimble-1996, Koh-Ong-1999, Lamarche-Strassburger-2006, Hughes-2012-nets}, each providing a different take on the proof equivalence problem for MLL. We will use the formulation by Hughes \cite{Hughes-2012-nets}, which reduces proof equivalence to a simple rewiring relation, as the basis for our PSPACE-hardness reduction. In this section we will discuss how this notion of proof net arises from the sequent calculus, and introduce a compact notation for it.

\begin{definition}[Proof nets]
\label{def:proof nets}
%
For a sequent $\Gamma$,
\begin{itemize}

	\item
	a \emph{linking} $\links$ is a function from the names of $\bot$-subformulae in $\Gamma$ to the names of $\Gamma$;

	\item
	a \emph{switching graph} for $\links$ is an undirected graph over the names of $\Gamma$ with:
	\begin{itemize}
		\item for every subformula $\named aA\named c\tn\named bB$ the edges $\edge ca$ and $\edge cb$,
		\item for every subformula $\named aA\named c\pr\named bB$ either the edge $\edge ca$ or the edge $\edge cb$, and
		\item for every subformula $\named a\bot$ the edge $\edge a{\links(a)}$;
	\end{itemize}

	\item
	a linking $\links$ is \emph{correct} if every switching graph is acyclic and connected;

 	\item
	a \emph{proof net} $(\Gamma,\links)$ consists of the sequent $\Gamma$ and a correct linking $\links$ for $\Gamma$.

@@

An example proof net is given in Figure~\ref{fig:example net}. The edges from a $\pr$-node, which are subject to being switched, are drawn as dashed lines, while edges from a $\tn$-node are solid lines. An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}, and may also be indicated $\jump a{\links(a)}$, to match the illustrations. Note that unlike the axiom links of proof nets without units, a jump may connect to any connective or unit, and multiple jumps may point to the same position.

\begin{figure}
\[
	\examplenet
\]
\caption{An example proof net}
\label{fig:example net}
\end{figure}

The present notion of proof net can be seen as a direct interpretation of the sequent calculus, in the following way. As there are inequivalent proofs of the sequent $\bot\tn\bot,1\pr1$, some way of attaching the $\bot$-formulae to the remainder of the proof net appears necessary. The introduction rule for $\bot$, in Figure~\ref{fig:MLL}, attaches a $\bot$-formula to a sequent---but sequents are not explicit in proof nets. The notion of a jump, which attaches the $\bot$ to an arbitrary point in the sequent $\Gamma$, is then a natural generalisation of the axiom links of unit-free MLL proof nets. This gives the following translation from proofs to proof nets.

\begin{definition}
\label{def:proofs to nets}
The relation $(\toNet)$ interprets a proof $\Pi$ by a linking $\links$ as follows: $\Pi\toNet\links$ if for each $\named a\bot$ in $\Pi$, if $\Gamma$ is the context of the inference introducing $\named a\bot$, as illustrated below, then $\links(a)$ is a name in $\Gamma$.
\[
	\infer[\MLLlabel b]{\Gamma,\named a\bot}{\Gamma}
\]
\end{definition}

\begin{proposition}[\cite{Danos-Regnier-1989,Lamarche-Strassburger-2006}]
\label{prop:correctness and sequentialisation}
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $(\Gamma,\links)$ is a proof net. For any proof net $(\Gamma,\links)$ there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (sequentialisation).
\end{proposition}

\begin{remark}
A main distinction with the proof nets of \cite{Blute-Cockett-Seely-Trimble-1994} is that the latter are defined as graphs whose vertices represent inferences, and whose edges represent formulae. This formulation allows an easy inclusion of cuts, which are not present in our proof nets. That the jumps from $\bot$-vertices in these graphs are attached to edges, not vertices, may seem curious at first, but it is a natural choice given the categorical view of $\bot$-introduction as the isomorphism $A\to A\pr\bot$.
\end{remark}

\subsection*{Rewiring}

The use of proof nets factors out the bottom three permutations in Figure~\ref{fig:permutations}, $(\pr-\pr$), $(\tn-\pr)$, and $(\tn-\tn)$. The remaining permutations, involving the $\bot$-introduction rule, impose an equivalence on proof nets, consisting of the \emph{rewiring} of jumps from one target to another. This is defined as follows.

\begin{definition}
\label{def:proof net equivalence}
A \emph{rewiring} $(\perm*)$ between proof nets changes the target of exactly one jump.
\emph{Equivalence} $(\perm)$ of proof nets over a sequent $\Gamma$ is the equivalence generated by rewiring.
\end{definition}

Figure~\ref{fig:rewiring} illustrates the rewiring relation in local steps. In the top row, a jump can be rewired across another jump and across a tensor without breaking correctness; these rewiring steps can be seen to correspond to $\bot-\bot$ and $\tn-\bot$ permutations. In the bottom row, rewiring a jump into the par, from $a$ or $b$, only preserves correctness if the jump can be rewired between $a$ and $b$ independently of the par. Intuitively, this corresponds to the $\pr-\bot$ permutation in the following way: the positions that a jump may rewire to correspond to the context of the rule introducing the $\bot$; then to have the premise of the $\pr$-introduction as its context, the jump should be able to rewire to both $a$ and $b$.

\begin{figure}
\[
\begin{array}{cccrccc}
	\PN b1 & \sim & \PN b2 &                          \PN t1 \quad \sim  & \PN t2 & \sim & \PN t3
\\ \\ \\
	\PN p1 & \sim & \PN p2 & \sim \quad \PN p3 \qquad\text{only if}\quad & \PN q1 & \sim & \PN q2
\end{array}
\]
\caption{Rewiring steps}
\label{fig:rewiring}
\end{figure}

A refinement of proof nets introduced by Hughes \cite{Hughes-2012-nets}, is that jumps may be restricted to target only $1$-formulae. For the remainder of the paper, unless otherwise indicated, we will assume that proof nets are of this form.

\begin{proposition}[\cite{Hughes-2012-nets}]
Any proof net is equivalent to one where the codomain of $\links$ is restricted to occurrences of the formula $1$.
\end{proposition}

A main technical advantage of this restriction is that cut-elimination can be performed in a one-shot operation, by path composition over jumps. Without the need to define cut-elimination by local transformations, cuts may then be omitted from proof nets altogether, which justifies the present definition.

Finally, the following proposition states that equivalence is preserved when translating between proofs and proof nets, which means that \capsabbrev{mll}\ proof equivalence is equivalent to the problem of deciding equivalence of proof nets.

\begin{proposition}[\cite{Blute-Cockett-Seely-Trimple-1994,Hughes-2012-nets}]
\label{prop:proof nets work}
%
If $\Pi\toNet\links$ and $\Pi'\toNet\links'$, then $\Pi\perm\Pi'$ if and only if $\links\perm\links'$.
%
\end{proposition}

@@

@@

We will use a concise diagrammatic notation for sequents and proof nets.
%
The units $1$ and $\bot$ are represented by a circle~($\circ$) and a disc~($\bullet$) respectively;
%
formulae related by a tensor will be connected by edges; 
formulae related by a par will be juxtaposed, and collected in a box when a par-formula is an immediate subformula of a tensor-formula (and also for illustrative purposes).
%
For example, the following denote the same sequent:
\[
	\bot\tn\bot,\bot\tn\bot,\1,(\1\pr\1\pr\1)\tn\bot
\]
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (2,0) node[bullet] (c) {} -- (3,0) node[bullet] (d) {};
	\node[circ] (1) at (4,0) {};
	\node[circ] (2) at (5.25,0) {}; \node[circ] (3) at (6.25,0) {}; \node[circ] (4) at (7.25,0) {};
	\draw[rounded corners] (4.75,-1) rectangle (7.75,1);
	\draw (7.75,0) -- (8.75,0) node[bullet] (e) {};
\end{tikzpicture}
\]
The links of a proof net are added to the sequent as coloured arrows.
%
The following example is a proof net for the above sequent.
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[circ] (1) at (6,0) {};
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (3,0) node[bullet] (c) {} -- (4,0) node[bullet] (d) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (2.center);
			\draw[bend right=10] (b) to (3.center);
			\draw[bend left=10]  (c) to (3.center);
			\draw[bend left=10]  (d) to (4.center);
			\draw[bend right=10] (e) to (1.center);
	\end{scope}
\end{tikzpicture}
\]

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\!\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\!\pr\cdots\pr\! A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
A path in a proof net or switching graph is indicated $\path ab$, and illustrated as below.
%
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\]

@@

As the linkings above illustrate, it may occur that one subformula of a weight element $A\pr B\pr C$ fills one half of a constraint element.
%
Consequently, a weight element can fill three halves, but only of different constraint elements.
%
The other three halves may be filled by weight elements of a different edge---so to fill 3 constraint elements requires 2 inappropriate edges.
%
To fill the next three constraint elements, at most 1 previous inappropriate edge may be used, and one additional one is needed.
%
To fill $3n$ constraint elements inappropriately therefore requires $n+1$ edges.
%
It thus suffices to multiply the number of constraint elements by three times the number of edges, and encode a vertex with inflow constraint $c$ by a vertex-gadget with $c\times3\times|E|$ constraint elements (where $|E|$ denotes the number of edges in the constraint graph).

@@

In the graph of a proof net, a chosen subnet for a par can be made explicit as a \emph{box}, as illustrated below left.
%
Boxes may replace the switching condition as a correctness criterion: in the example, both the outside and the inside of the box form a tree.
%
To make this precise, we will consider the action of \emph{closing} a box, which means it is regarded as a single vertex in the graph, as illustrated below right.
%
%
%
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {}
		  (.25,0) node[bullet] (a) {} -- (1.25,0) node[bullet] (b) {}
		  ( 2.75,0) node[bullet] (c) {} -- (3.75,0) node[bullet] (d) {};
	\node[circ] (0) at (5.75,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5.5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (2);
			\draw[bend right=10] (a) to (2);
			\draw[bend right=10] (b) to (3);
			\draw[bend left=10]  (c) to (3);
			\draw[bend left=10]  (d) to (4);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\draw[dashed,rounded corners] (-.25,-.5) rectangle (4.25,4.25);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {};
	\node[circ] (0) at (3,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[draw,rounded corners,minimum size=16pt] (box) at (1,3) {};
	\draw (box)--(3,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (box);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\path (0,4.25)--(1,4.25);
\end{tikzpicture}
\]
%
%
\begin{definition}
A \emph{boxing} $s$ for a linking $\links$ for $\Gamma$ assigns a sub-sequent $s(v)\leq\Gamma$ to each par-vertex $\named v\pr$ such that 1) $v$ is a port of $s(v)$ and 2) boxes are either disjoint or strictly nested: if $s(v)\cap s(w)\neq\varnothing$ then $s(w)<s(v)$ or $s(v)<s(w)$.
\end{definition}

@@

The linearity of \capsabbrev{mll} means that in a proof or proof net, there is always a certain balance to the number of $\bot$- and $\pr$-occurrences. 
%
This observation gives a well-known necessary condition for the provability of a sequent.

\begin{definition}
The \emph{balance} of a sequent is the number of $\bot$s minus the number of $\pr$s and commas.
%
A sequent is \emph{balanced} if its balance is zero.
\end{definition}
